---
title: "SynergyScreen Quick Start"
author: "Yury V Bukhman"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette: 
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{SynergyScreen Quick Start}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))

# SynergyScreen Quick Start

```{r attach SynergyScreen}
library(SynergyScreen)
```

## Generate design for a new screen
The starting point is a CSV file listing compounds to be used in the screen. It should have 3 columns: compound, min.dose, and	max.dose. Here is an example:
```{r see the compounds file}
read.csv(system.file("extdata/8_compounds/8_compounds.csv", package = "SynergyScreen"))
```

Load compounds into a special object that will be used to initialize a new screen
```{r Load a set of compounds}
compounds <- readCompoundFile(system.file("extdata/8_compounds/8_compounds.csv", package = "SynergyScreen"))
```

Initialize a new screen
```{r Initialize a new screen}
screen0 <- new("SynergyScreen", compound_list=compounds)
```

Generate screen design for 96-well plates
```{r Generate screen design for 96-well plates}
screen0 <- generateDesign(screen0, type="12 doses exp") 
write.csv(design(screen0), file="8_compounds_design.csv", row.names=F)
```

## Analyze screen data
First, load a design file^[The difference between the auto-generated design and the design that was actually used was that controls were in rows G and H in the first and A and H in the second.] and a file with screening data, e.g. ODs of plate wells.
```{r Read in the data}
screen1 <- designFile2SynergyScreen(system.file("extdata/8_compounds/8_compounds_design_actually_used.csv", package = "SynergyScreen")) 
OD.data <- readWideDataFlex(file = system.file("extdata/8_compounds/8_compounds_OD_data.csv", package = "SynergyScreen"), row = c(3,13,23,33,43,53), col = rep("G",6)) 
raw_data(screen1) <- OD.data 
```

Normalize and compute synergies 
```{r Normalize and compute synergies}
screen1 <- normalize(screen1) 
screen1 <- computeSynergies(screen1) 
```

Find synergies and antagonisms^[menls stands for "median effect non-linear least squares". It fits dose-response data using the median effect equation. Other options include "melm" and "loess". See vignette "Introduction to SynergyScreen" for details.]
```{r Find synergies and antagonisms}
findSynergies(screen1, statistic="max", threshold=0.7, direction="lower") 
findSynergies(screen1, statistic="min", threshold=1.5, direction="higher") 
```

## Visualize screen data 
Plot a dose-response experiment 
```{r Plot a dose-response experiment}
plot(dre(screen1,"Cpd_2")) 
```

Plot a isoboles for a pair of compounds 
```{r Plot a isoboles for a pair of compounds}
plotIsoboles(screen1, "Cpd_1-Cpd_4", effect=seq(0.1,0.7,by=0.1)) 
```

Plot a heat map of the interaction matrix 
```{r Plot a heat map of the interaction matrix}
plot(screen1) 
```
