---
title: "SynergyScreen Quick Start"
author: "Yury V Bukhman"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette: 
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{SynergyScreen Quick Start}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo=FALSE}
knitr::opts_chunk$set(fig.height=4, fig.width=6)
```

```{r attach SynergyScreen, message=FALSE}
library(SynergyScreen)
```

## Generate design for a new screen
The starting point is a CSV file listing compounds to be used in the screen. It should have 3 columns: compound, min.dose, and	max.dose. Here is an example:
```{r see the compounds file}
read.csv(system.file("extdata/8_compounds/8_compounds.csv", package = "SynergyScreen"))
```

Load compounds into a special object that will be used to initialize a new screen
```{r Load a set of compounds}
compounds <- readCompoundFile(system.file("extdata/8_compounds/8_compounds.csv", package = "SynergyScreen"))
```

Initialize a new screen
```{r Initialize a new screen}
screen0 <- new("SynergyScreen", compound_list=compounds)
```

Generate screen design for 96-well plates
```{r Generate screen design for 96-well plates}
screen0 <- generateDesign(screen0, type="12 doses exp") 
write.csv(design(screen0), file="8_compounds_design.csv", row.names=F)
```
Ignore the warning given by `write.csv()`

Display dose-response experiments, their positions on the plate, maximal doses and serial dilution factors. 
```{r designSummary}
designSummary(screen0)
```
This helps the experimenter to set up the screen. The doses are in exponential series. A lower compound concentration is obtained by adding a specified amount of blank medium to a medium that has a higher concentration of the same compound. For example, if dilution factor = 0.1, get a lower concentration by adding 0.1 volume of blank medium to 1 volume of a medium with a higher concentration.

## Analyze screen data

First, load a design file^[The difference between the auto-generated design and the design that was actually used was that controls were in rows G and H in the first and A and H in the second.] and a file with screening data, e.g. ODs of plate wells.
```{r Read in the data}
screen1 <- designFile2SynergyScreen(system.file("extdata/8_compounds/8_compounds_design_actually_used.csv", package = "SynergyScreen")) 
OD.data <- readWideDataFlex(file = system.file("extdata/8_compounds/8_compounds_OD_data.csv", package = "SynergyScreen"), row = c(3,13,23,33,43,53), col = rep("G",6)) 
raw_data(screen1) <- OD.data 
```

Normalize and compute synergies 
```{r Normalize and compute synergies}
screen1 <- normalize(screen1) 
screen1 <- computeSynergies(screen1) 
```

Find synergies and antagonisms^[menls stands for "median effect non-linear least squares". It fits dose-response data using the median effect equation. Other options include "melm" and "loess". See vignette "Introduction to SynergyScreen" for details.]
```{r Find synergies and antagonisms}
findSynergies(screen1, statistic="max", threshold=0.7, direction="lower") 
findSynergies(screen1, statistic="min", threshold=1.5, direction="higher") 
```

## Visualize screen data 
Plot a heat map of the interaction matrix 
```{r Plot a heat map of the interaction matrix, warning=FALSE}
plot(screen1) 
```

Note that the heat map does not show compounds 5, 6, and 7. This is because we do not have data for these compounds at 50% inhibition. 

Plot isoboles of synergistic compounds
```{r Plot isoboles of synergistic compounds}
plotIsoboles(screen1, "Cpd_3-Cpd_4")
```

Plot isoboles of antagonistic compounds
```{r Plot isoboles of antagonistic compounds}
plotIsoboles(screen1, "Cpd_1-Cpd_4", effect=seq(0.1,0.7,by=0.1)) 
```

Plot isoboles of non-interacting compounds
```{r Plot isoboles of non-interacting compounds}
plotIsoboles(screen1, "Cpd_2-Cpd_3")
```

Plot the dose-response curve of a compound that worked. Note the sigmoid shape, well modeled by the median effect equation.
```{r Plot the dose-response curve of a compound that worked}
plot(dre(screen1,"Cpd_2")) 
```

Plot the dose-response curve of a compound that did not work. Note the narrow response range. Next time, the experimenter may want to try a much wider dose range.
```{r Plot the dose-response curve of a compound that did not work}
plot(dre(screen1,"Cpd_6")) 
```

